### 2014-03-18 HowtoHowtoCapon/submission/submit1/answer/answer.tex
la2pdf commit - 2014:03:18	13:21:37	(Tue18.Mar)

\begin{enumerate}
\qi{It is unfortunate that so little effort was made to speed up the CPU versions of the code. For example, there are public-domain versions of optimized LAPACK libraries that provide the same sort of linear algebra codes as the authors used in their CPU implementation (for example, GoToBLAS, or Intel's MKL libraries).}
\ai{Thank you. We have now included our reasons for not comparing with an optimized CPU implementaiton.  One reason is that we have not been able to find any CPU batch solvers out there and lacked the time to make our own. In our tests our custom inline Cholesky solver beat alternatives from e.g. MKL as our matrices are too small to justify the extra function call overhead. Another reason is that to make the comparison fair we feel an equal effort would have to put into writing a covariance builder for the CPU, and to minimize data movement in each step and in between them. }
%
\qi{The extra input/output of data to the GPU is a burden only on the GPU implementation, so would penalize the GPU implementation only, unless the authors were to "hide" the i/o by setting it up to run in parallel with the CPU processing, which is easier and easier with the latest GPU architectures. This should perhaps be pointed out.}
\ai{Yes this is a good point. This is now included.}
%
\qi{The authors do not provide much information about the GPU hardware they used, the NVIDIA Quadro 6000. How does this compare to the desktop video cards that reside in most desktop, workstation, server and even high-end laptop systems? What is the vintage of the architecture of this GPU? How has the architecture evolved? For example, the authors tell us the Quadro 6000 (\$1900, 6GB VRAM, 448 cores) is a Fermi architecture - this is the previous architecture, having been superceded by the Kepler architecture, with a new Quadro K6000 board (\$3600, 12 GB VRAM, 2880 cores). The Quadro boards are high-end boards with video output targeting the CADAM market - there are many cheaper alternatives tar-geted at gangers and an even more expensive compute-only board called TESLA targeted at high performance computing. It's difficult to say how much of this irrelevant. }
\ai{TODO}
%
\qi{The authors state that OpenCL is an equally attractive alternative to CUDA. This would be the case if the authors had been able to find something in OpenCL akin to the recently released batch-optimized linear solver code for CUDA. Unfortunately, this is a pervasive story - the support for scientific programming in CUDA is years ahead of OpenCL. The benefit of OpenCL is that it is supposed to be "cross-platform", which is true in theory, but it is hard to imagine being true in practice considering the wide range of heterogeneous computing elements OpenCL is supposed to cover; from phones to DSP chips to FPGAs. }
\ai{These are important remarks, thank you. We have written OpenCL programs too, and run them on Intel CPUs and GPUs from AMD and Nvidia. It was no walk in the park, but with a satisfactory end result. But like you say, these are just a subset of devices that OpenCL support, and we have now changed the wording to be more specific here. Your remark on the lack a batched linear solver is spot in, the sentence is now rewritten.}
\end{enumerate}

>>>>

\begin{enumerate}
\qi{It is unfortunate that so little effort was made to speed up the CPU versions of the code. For example, there are public-domain versions of optimized LAPACK libraries that provide the same sort of linear algebra codes as the authors used in their CPU implementation (for example, GoToBLAS, or Intel's MKL libraries).}
\ai{Thank you. We have now included our reasons for not comparing with an optimized CPU implementation.  One reason is that we have not been able to find any CPU batch solvers out there and lacked the time to make our own. In our tests our custom inline Cholesky solver beat alternatives from e.g. MKL as our matrices are too small to justify the extra function call overhead. Another reason is that to make the comparison fair we feel an equal effort would have to put into writing a covariance builder for the CPU, and to minimize data movement in each step and in between them. The details may be found in section V paragraph 2.}
%
\qi{The extra input/output of data to the GPU is a burden only on the GPU implementation, so would penalize the GPU implementation only, unless the authors were to "hide" the i/o by setting it up to run in parallel with the CPU processing, which is easier and easier with the latest GPU architectures. This should perhaps be pointed out.}
\ai{Yes this is a good point. Included in last paragraph of section IV.}
%
\qi{The authors do not provide much information about the GPU hardware they used, the NVIDIA Quadro 6000. How does this compare to the desktop video cards that reside in most desktop, workstation, server and even high-end laptop systems? What is the vintage of the architecture of this GPU? How has the architecture evolved? For example, the authors tell us the Quadro 6000 (\$1900, 6GB VRAM, 448 cores) is a Fermi architecture - this is the previous architecture, having been superceded by the Kepler architecture, with a new Quadro K6000 board (\$3600, 12 GB VRAM, 2880 cores). The Quadro boards are high-end boards with video output targeting the CADAM market - there are many cheaper alternatives targeted at gamers and an even more expensive compute-only board called TESLA targeted at high performance computing. It's difficult to say how much of this irrelevant.}
\ai{There is certainly a potential for comparing with many other cards here, but we decided to be thorough with one architecture than to compare with everything. To the paragraph introducing the Quadro 6000 we have explained what its equivalent consumer grade Geforce card would be. We have also included remarks that we expect Kepler to perform better had the design been optimized for it.}
%
\qi{The authors state that OpenCL is an equally attractive alternative to CUDA. This would be the case if the authors had been able to find something in OpenCL akin to the recently released batch-optimized linear solver code for CUDA. Unfortunately, this is a pervasive story - the support for scientific programming in CUDA is years ahead of OpenCL. The benefit of OpenCL is that it is supposed to be "cross-platform", which is true in theory, but it is hard to imagine being true in practice considering the wide range of heterogeneous computing elements OpenCL is supposed to cover; from phones to DSP chips to FPGAs. }
\ai{These are important remarks, thank you. We have written OpenCL programs too, and run them on Intel CPUs and GPUs from AMD and Nvidia. It was no walk in the park, but with a satisfactory end result. But like you say, these are just a subset of devices that OpenCL support, and we have now changed the wording to be more specific here. Your remark on the lack a batched linear solver is spot in, the sentence is now rewritten.}
\end{enumerate}
